Cooling Requirements for the Vertical Shear Instability, \\ with Implications for Protoplanetary Disks

### 2015-04-08 appendix.tex
moving more equations to appendix
restructured: moved more equations to appendix, re-ordered appendix, slightly shortened analytical section

\section{Applicability of the vertically global shearing box}\label{global_corr}
In \S\ref{analytical} we effectively employed the vertically global shearing box
\citep[VGSB,][]{mcnally14} to make analytic progress.   
The difference between the linear problem in the VGSB 
and the radially-local approximation adopted in our numerical study
is the absence of terms explicitly associated with the radial disk structure in
the VGSB. These terms appear in the energy, mass and radial momentum
equations and are labeled by $\hat{g}_c$ in
Eq. \ref{lin_mass}---\ref{lin_energy}. Nominally $\hat{g}_c=1$.

@@

However, in the VGSB these terms are neglected by 
setting $\hat{g}_c=0$. For a power-law disk, these radial gradients
are $O(r^{-1})$, and they appear in comparison with terms of
$O(k_x)$. The neglected terms therefore have a relative magnitude of
$O(\epsilon/\khat)$. This is small for thin disks ($\epsilon\ll1$)
and/or small radial wavelengths ($\khat\gg 1$).

\subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
A limitation of the VGSB is that it cannot be used in the
adiabatic limit, even if the above conditions are satisfied. We explain this by considering vertically isothermal
disks and setting  $\beta\to\infty$ in 
Eq. \ref{vertiso_gov_nondim} (which itself is obtained by setting
$\hat{g}_c=0$ in Eq. \ref{lin_mass}---\ref{lin_energy}) to give 
\begin{align}
  0 =\dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime
  +\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
    -\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
\end{align}
We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
integrate vertically, assume boundary terms vanish when integrating by
parts, to obtain
\begin{align}
  \hat{\sigma}^2\left(\frac{1}{\gamma} +
    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
  v_z|^2 d\zhat 
  =&  \left(\frac{\gamma-1}{\gamma}\right)
  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
  v_z)^\prime|^2 d\zhat\notag\\
&+
  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
    2}|\delta v_z|^2 d\zhat
+ \ii\epsilon q \hat{k}
  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
  \delta v_z d\zhat.\label{adia_integral}
\end{align}
% When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
% $\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
% expected, a sub-adiabatically stratified disk is stable in the absence
% of vertical shear. 
In the presence of vertical shear $q\neq0$, Eq. \ref{adia_integral}
suggests $\hat{\sigma}^2$ is complex for real $\khat$, which implies the
possibility of instability, regardless of the value of
$|\gamma-1|$. This is contradicts the second Solberg-Hoiland criterion 
(Eq. \ref{solberg2}), which imply instability for adiabatic
perturbations in a thin, vertically isothermal disk would require 
$|\gamma-1|\ll1$. 
%also need real freq^2 in adia limit 
This is because the VGSB retains the vertical shear
($\p_z\Omega$) from the global disk, but ignores the global radial
disk structure responsible for it. Nevertheless, we demonstrate below
that this inconsistency is unimportant for the VSI because it occurs
for $\beta\ll1$.   %and small scale

@@

\section{Full governing equation for vertically isothermal disks in
  the VGSB}\label{adia_improve}

In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
eliminating variables to obtain a single equation for $\delta v_z$ in
vertically isothermal disks, Eq. \ref{vertiso_gov}.  
This procedure ignores the vertical dependence of
$D=\kappa^2(z) - \sigma^2$. We show here that this has no serious 
consequence for thin disks.

In the global disk with $\Gamma=1$ we have
\begin{align}\label{dkappa2}
  \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
    r}\frac{\p\Omega^2}{\p z} = -
   \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(
  \frac{3z^2}{r^2+z^2}-1\right) \equiv - \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2}F(r,z).
\end{align}
% For the local problem we may then write
% \begin{align}
%   \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z),
% \end{align}
The function $F$ increases monotonically from $F=-1$ at $z=0$ to $F\to2$
as $|z|\to\infty $, so $|F|=O(1)$. 
% For the local problem we set
% $dD/dz = \p\kappa^2/\p z$. 
% This function varies from $F=q$ at $z=0$ to $F\to
% 3+q$ as $|z|\to\infty$. Since we consider $|q|=O(1)$, this impies
% $|F|=O(1)$ as well.

For the radially local problem, we eliminate $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
Eq. \ref{lin_vz}  (with $\Gamma=1$), but now retain the vertical dependence
of $D$ using Eq. \ref{dkappa2}, to obtain  
\begin{align}
  0 =& \frac{d^2\delta v_z}{dz^2} + \left[1 + \frac{\ii k_x c_s^2
      q}{Dr} - \frac{k_x^2c_s^2}{\left(k_x^2c_s^2 + \chi
        D\right)}\frac{qc_s^2F}{Dr^2}\right]\frac{d\ln\rho}{dz}\frac{d\delta
    v_z}{dz} \notag\\
  &+ \left\{\sigma^2\left(\frac{k_x^2}{D} +
      \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
        q}{Dr}\right)\frac{d^2\ln\rho}{dz^2} -
    \frac{c_s^2}{D}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
      \frac{\ii k_x q}{r}\right)
   \left[\left(1-\chi\right) +
     \frac{\chi}{\left(k_x^2c_s^2 + \chi D\right)}\frac{qc_s^2 F}{r^2}\right] 
   \right\}\delta v_z,
\end{align}
where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
t_c \gamma\right)$. Making the low frequency, Keplerian
approximations with the replacement $D\to \Omega_k^2$ gives, in terms
of dimensionless variables,
\begin{align}
   &\delta v_z ^{\prime\prime} + \left[1 + \ii\epsilon q\hat{k} -
    \frac{ \hat{k}^2}{
      \left(\hat{k}^2+\chi\right)}q \epsilon^2F\right]\ln\rho^{\prime}\delta v_z^\prime +
  \left\{\left(\chi + \ii \epsilon q
      \hat{k}\right)\ln\rho^{\prime\prime} - \ln\rho^{\prime
      2}\left(\khat^2 -
      \ii\epsilon
      q\hat{k}\right)\left[1 - \chi +
      \frac{\chi}{\left(\hat{k}^2+\chi\right)}q\epsilon^2F\right]\right\}\delta v_z \notag\\&=
  -\hat{\sigma}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
\end{align}   
Eq. \ref{adia_diso3} differs from Eq. \ref{vertiso_gov_nondim} by terms
proportional to $\epsilon^2$. For a thin disk ($\epsilon\ll1$) with
$|q|=O(1)$, as considered in this work, neglecting these terms has no
qualitative effect on the discussion in \S\ref{analytic_relax}.

>>>>

\section{Full governing equation for vertically isothermal disks in
  the vertically global shearing box}\label{adia_improve}
Here we explain the derivation of Eq. \ref{vertiso_gov} used in our 
analytical discussion in \S\ref{analytical}. We shall need the
vertical derivative of the epicyclic frequency $\kappa$. In the global
disk with $\Gamma=1$ we have 
\begin{align}\label{dkappa2}
  \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
    r}\frac{\p\Omega^2}{\p z} = -
  \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(
    \frac{3z^2}{r^2+z^2}-1\right) \equiv - \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2}F(r,z).
\end{align}
% For the local problem we may then write
% \begin{align}
%   \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z),
% \end{align}
The function $F$ increases monotonically from $F=-1$ at $z=0$ to $F\to2$
as $|z|\to\infty $, so $|F|=O(1)$.

In the vertically global shearing box (VGSB) approximation, we set
$\hat{g}_c=0$ in Eq. \ref{lin_mass} --- \ref{lin_energy}. We then eliminate
the horizonatal velocity perturbations ($\delta v_x,\, \delta v_y$) to
obtain  
%we can eliminate variables 
%to obtain a pair or ordinary differential
%equations for $W,Q$ as 
\begin{align}
  & \ii\sigma\delta v_z = \frac{dW}{dz} + \frac{d\ln{\rho}}{d z}\left(W-Q\right),\label{ode_vz} \\
  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W =\ii\sigma \left(\frac{\ii
      k_x r}{D}\frac{d\Omega^2}{dz} -
    \frac{d\ln{\rho}}{dz}\right)\delta v_z - \ii\sigma \frac{d\delta v_z}{dz},
  % \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] 
  % \notag\\
  % &- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
  %   \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),
  \label{ode_w}\\
%  &\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
%  \frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right) 
%  =\ii \sigma c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} -
%    1\right)\delta v_z,  
  &\sigma^2W - \gamma\sigma^2 Q +
  \frac{\ii\sigma}{t_c}\left(W-Q\right) 
  =\ii \sigma c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma -
    1\right)\delta v_z,
  % 
  % \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],
  \label{ode_Q} 
\end{align}
(note that $\Gamma=1$ has been used) where
\begin{align}
  D \equiv \kappa^2 - \sigma^2,
\end{align} 
and we have replaced the notation $\p_z$ by $d/dz$ since we are
considering solutions at a fixed radius.

% In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
% eliminating variables to obtain a single equation for $\delta v_z$ in
% vertically isothermal disks, Eq. \ref{vertiso_gov}.  
% This procedure ignores the vertical dependence of
% $D=\kappa^2(z) - \sigma^2$. We show here that this has no serious 
% consequence for thin disks.

%For the radially local problem, we eliminate $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
%Eq. \ref{lin_vz}  (with $\Gamma=1$), but now retain the vertical dependence
%of $D$ using Eq. \ref{dkappa2}, to obtain

We now eliminate $W,\, Q$ from Eq. \ref{ode_vz} --- \ref{ode_Q} in
favor of $\delta v_z$, making use of Eq. \ref{dkappa2} to evaluate
$dD/dz$, to obtain

\begin{align}\label{full_ode}
  0 =& \frac{d^2\delta v_z}{dz^2} + \left[1 + \frac{\ii k_x c_s^2
      q}{Dr} - \frac{k_x^2c_s^2}{\left(k_x^2c_s^2 + \chi
        D\right)}\frac{qc_s^2F}{Dr^2}\right]\frac{d\ln\rho}{dz}\frac{d\delta
    v_z}{dz} \notag\\
  &+ \left\{\sigma^2\left(\frac{k_x^2}{D} +
      \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
        q}{Dr}\right)\frac{d^2\ln\rho}{dz^2} -
    \frac{c_s^2}{D}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
      \frac{\ii k_x q}{r}\right)
   \left[\left(1-\chi\right) +
     \frac{\chi}{\left(k_x^2c_s^2 + \chi D\right)}\frac{qc_s^2 F}{r^2}\right] 
   \right\}\delta v_z,
\end{align}
where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
t_c \gamma\right)$. Making the low frequency and Keplerian
approximations is equivalent to the replacement $D\to
\Omega_k^2$. This gives, in terms of dimensionless variables,  
\begin{align}
  &\delta v_z ^{\prime\prime} + \left[1 + \ii\epsilon q\hat{k} -
    \frac{ \hat{k}^2}{
      \left(\hat{k}^2+\chi\right)}q \epsilon^2F\right]\ln\rho^{\prime}\delta v_z^\prime +
  \left\{\left(\chi + \ii \epsilon q
      \hat{k}\right)\ln\rho^{\prime\prime} - \ln\rho^{\prime
      2}\left(\khat^2 -
      \ii\epsilon
      q\hat{k}\right)\left[1 - \chi +
      \frac{\chi}{\left(\hat{k}^2+\chi\right)}q\epsilon^2F\right]\right\}\delta v_z \notag\\&=
  -\hat{\sigma}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
\end{align} 
Since we are interested in thin disks with $\epsilon \ll 1$ and $|q| =
O(1)$, we neglect the terms proportional to $\epsilon^2$, which gives 
Eq. \ref{vertiso_gov_nondim} (the dimensional version being
Eq. \ref{vertiso_gov}). This is equivalent to neglecting the   
vertical dependence of $D$ by setting $D\to\Omega_k^2$ in
Eq. \ref{ode_w} before eliminating variables in favor of $\delta
v_z$, which is the simpler way to obtain Eq. \ref{vertiso_gov}.

%Thus, a simpler route to the governing
%equation (Eq. )

%Eq. \ref{adia_diso3} differs from Eq. \ref{vertiso_gov_nondim} by terms
%proportional to $\epsilon^2$. For a thin disk ($\epsilon\ll1$) with
%$|q|=O(1)$, as considered in this work, neglecting these terms has no
%qualitative effect on the discussion in \S\ref{analytic_relax}. 
% In deriving Eq. \ref{vertiso_gov}, the vertical dependence of
% $D=\kappa^2(z)-\sigma^2$ was ignored because the replacement
% $D\to\Omega_k^2$ was made before eliminating variables in favor of
% $\delta v_z$. In Appendix \ref{adia_improve} we show that making this
% replacement after eliminating variables, which involves $dD/dz$,
% introduces terms of $O(\epsilon^2)$. Since we are interested in 
% thin disks ($\epsilon\ll 1 $), e.g. protoplanetary disks where 
% $\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce
% unnecessary complexity for the present discussion.

\section{Applicability of the  VGSB}\label{global_corr}
%In \S\ref{analytical} we effectively employed the vertically global
%shearing box \citep[VGSB,][]{mcnally14} to make analytic progress. 
The vertically global shearing box \citep[VGSB,][]{mcnally14} is an
extension of the standard shearing box \citep[SSB,][]{goldreich65} to
background shear flows that are height dependent. 
%As in the SSB, in
%the VGSB the equilibrium disk structure is assumed to be homogeneous
%in the radial ($x$) direction.  
Within the VGSB framework, the vertical shear $d\Omega/dz$
can be freely specified. We choose $d\Omega/dz$ to be that due to the
radial temperature gradient in the global disk
(Eq. \ref{vertical_shear}). However, explicit dependencies on the
radial disk structure are ignored elsewhere in the linearized equations
as described below. We show in the following sections that this
inconsistency only becomes important in the adiabatic limit, which is
not the relevant regime for the VSI.

%is only important when
%considering adiabatic limit, which is not the relevant regime for the
%VSI.

The background radial disk structure explicitly appears 
in the linearized mass, radial momentum and energy equations
(Eq. \ref{lin_mass}, \ref{lin_vx} and \ref{lin_energy}) as the terms
proportional to $\hat{g}_c$. Nominally $\hat{g}_c=1$.      
In the VGSB (and SSB) these terms are neglected by setting
$\hat{g}_c=0$. For a power-law disk, these radial gradients  
are $O(r^{-1})$, and they appear in comparison with terms of
$O(k_x)$. The neglected terms therefore have a relative magnitude of
$O(\epsilon/\khat)$, which is small for thin disks ($\epsilon\ll1$)
and/or small radial wavelengths ($\khat\gg 1$).

% The difference between the linear problem in the VGSB 
% and the radially-local approximation adopted in our numerical study
% is the absence of terms explicitly associated with the radial disk structure in
% the VGSB.

%However, background radial gradient terms 
%are ignored elsewhere in the VGSB (which is done by setting
%$\hat{g}_c=0$).

@@

\subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
A limitation of the VGSB is that it cannot be used in the
adiabatic limit, even if the above conditions are satisfied. We explain this by considering vertically isothermal
disks and setting  $\beta\to\infty$ in 
Eq. \ref{vertiso_gov_nondim} % (which itself is obtained by setting
% $\hat{g}_c=0$ in Eq. \ref{lin_mass}---\ref{lin_energy})
to give 
\begin{align}
  0 =\dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime
  +\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
    -\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
\end{align}
We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
integrate vertically, assume boundary terms vanish when integrating by
parts, to obtain
\begin{align}
  \hat{\sigma}^2\left(\frac{1}{\gamma} +
    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
  v_z|^2 d\zhat 
  =&  \left(\frac{\gamma-1}{\gamma}\right)
  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
  v_z)^\prime|^2 d\zhat\notag\\
&+
  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
    2}|\delta v_z|^2 d\zhat
+ \ii\epsilon q \hat{k}
  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
  \delta v_z d\zhat.\label{adia_integral}
\end{align}
% When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
% $\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
% expected, a sub-adiabatically stratified disk is stable in the absence
% of vertical shear. 
In the presence of vertical shear $q\neq0$, Eq. \ref{adia_integral}
suggests $\hat{\sigma}^2$ is complex for real $\khat$, which implies the
possibility of instability, regardless of the value of
$|\gamma-1|$. This is contradicts the second Solberg-Hoiland criterion 
(Eq. \ref{solberg2}), which imply instability for adiabatic
perturbations in a thin, vertically isothermal disk would require 
$|\gamma-1|\ll1$. 
%also need real freq^2 in adia limit 
This is because the VGSB retains the vertical shear
($\p_z\Omega$) from the global disk, but ignores the global radial
disk structure responsible for it. Nevertheless, we demonstrate below
that this inconsistency is unimportant for the VSI because it occurs
for $\beta\ll1$.   %and small scale

@@

### 2015-04-08 linear_iso.tex
restructured: moved more equations to appendix, re-ordered appendix, slightly shortened analytical section

In Appendix \ref{adia_improve}, we derive a governing equation with the following approximations:

@@

\item%  We adopt the  `vertically global shearing box' (VGSB) formalism
%   developed by \cite{mcnally14}, which is an extension to the standard
%   shearing box \citep{goldreich65} background shear flows that are
%   height dependent.  As in the standard shearing box, in the VGSB the equilibrium disk
%   structure is assumed to be homogeneous in the radial ($x$) direction.

% % Thus, the linear problem in the VGSB can be derived from
% % Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
% % $\hat{g}_c=0$.

% in which explicit dependences on the background radial disk structure
%   in Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
%   $\hat{g}_c=0$.

% However, the global radial temperature structure,
%   which gives rise to vertical shear, is retained.

\item We consider eigenvalues such that
  $|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
  \kappa^2$. %\citep[e.g.][]{lin10}. 
  We call this the low frequency approximation. We are therefore
  considering inertial waves \citep{barker15}.

@@

%The result for $\Gamma=1$ is
\begin{align}
  0 =& \frac{d^2\delta v_z}{dz^2} + \left(1 + \frac{\ii k_x c_s^2
      q}{\Omega_k^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
    v_z}{dz} \notag\\
  % &+ \left[\sigma^2\left(\frac{k_x^2}{\Omega_k^2} +
  %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
  &+ \left[\frac{\sigma^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_k^2} +
      \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
        q}{\Omega_k^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
  &\phantom{+=}\left.-
    \frac{c_s^2}{\Omega_k^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
      \frac{\ii k_x q}{r}\right)
    \left(1-\chi\right) 
  \right]\delta v_z,\label{vertiso_gov}
\end{align}
where
\begin{align}
  \chi \equiv \frac{1-\ii\sigma t_c}{1-\ii\sigma t_c\gamma}.
\end{align}

@@

In deriving Eq. \ref{vertiso_gov}, the vertical dependence of
$D=\kappa^2(z)-\sigma^2$ was ignored because the replacement
$D\to\Omega_k^2$ was made before eliminating variables in favor of
$\delta v_z$. In Appendix \ref{adia_improve} we show that making this
replacement after eliminating variables, which involves $dD/dz$,
introduces terms of $O(\epsilon^2)$. Since we are interested in 
thin disks ($\epsilon\ll 1 $), e.g. protoplanetary disks where 
$\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce
unnecessary complexity for the present discussion.

>>>>

In Appendix \ref{adia_improve} we derive a governing equation with the following simplifications:

@@

\item  We adopt the  `vertically global shearing box' (VGSB) formalism 
  developed by \cite{mcnally14}, in which explicit dependencies on the
  background radial disk structure in 
  Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
  $\hat{g}_c=0$. However, the global radial temperature 
  gradient in the expression for the 
  vertical shear (Eq. \ref{vertical_shear}) is retained.

  In Appendix \ref{global_corr}  we discuss limitations of using
  the VGSB, where we show that this approximation is valid provided we 
  consider consider fast thermal relaxation times
  ($\beta\lesssim 1$) and perturbations with small radial lengthscales
  ($k_xH\gtrsim 1$), as appropriate for the VSI.

\item We consider eigenvalues such that
  $|\sigma^2|\ll \kappa^2,\,\Omega^2$. We call this the low frequency
  approximation. We are therefore considering inertial waves
  \citep{barker15}.

@@

These approximations lead to Eq. \ref{full_ode} which, for a thin disk
$\epsilon\ll 1$, simplifies further to  
%The result for $\Gamma=1$ is
\begin{align}
  0 =& \frac{d^2\delta v_z}{dz^2} + \left(1 + \frac{\ii k_x c_s^2
      q}{\Omega_k^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
    v_z}{dz} \notag\\
  % &+ \left[\sigma^2\left(\frac{k_x^2}{\Omega_k^2} +
  %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
  &+ \left[\frac{\sigma^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_k^2} +
      \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
        q}{\Omega_k^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
  &\phantom{+=}\left.-
    \frac{c_s^2}{\Omega_k^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
      \frac{\ii k_x q}{r}\right)
    \left(1-\chi\right) 
  \right]\delta v_z,\label{vertiso_gov}
\end{align}
where
\begin{align}
  \chi \equiv \frac{1-\ii\sigma t_c}{1-\ii\sigma t_c\gamma}.
\end{align}

@@
